---
title: "Normal Distribution Practice"
subtitle: "MATH 361"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

Let's simulate 100,000 SAT scores, which have a normal distribution with mean 1500 and standard deviation 200. We'll then convert them to Z-scores to explore the properties of Z-scores and the normal distribution more broadly.

Some questions to explore:

-   What is the shape of both distributions?
-   What is the range of both distributions?
-   What is the center of both distributions?
-   What makes some z-scores negative?
-   Do larger distances from the mean correspond to smaller or larger z-scores?
-   What % of values fall within 1 sd of the mean?
-   What % of values fall within 2 sds of the mean?
-   What % of values fall within 3 sds of the mean?
-   What changes about any of the above properties when you change the mean and the standard deviation?

```{r}

```

## Practice with `pnorm()`

### Task 1

Assume $Z \sim N(0,1)$. Find the following:

a.  $P(Z \leq 1.25)$
b.  $P(1.25 \leq Z \leq 2.31)$
c.  $P(-2.31 \leq Z \leq -1.25)$
d.  $P(Z \leq -2.31)$
e.  $P(-2.31 \leq Z \leq 1.25)$
f.  $P(Z > 1.25)$

```{r}

```

If you haven't already, draw pictures of each of the probabilities you computed above. How do b & c relate to one another?

> YOUR ANSWER HERE

What about a, d, & e?

> YOUR ANSWER HERE

What about a & f?

> YOUR ANSWER HERE

## Task 2

The textbook defines a quantity $z_{\alpha}$ as: $P(Z \geq z_{\alpha}) = \alpha$. (This is unusual as it is the $(1-\alpha)th$ percentile). Find:

a.  $z_{0.9147}$
b.  $z_{0.0125}$
c.  $z_{0.05}$
d.  $-z_{0.025}$

```{r}

```

### Task 3

Let $X \sim N(3,16)$. Find $P(4 \leq X \leq 8)$.

```{r}

```

### Task 4

*Note, this problem comes from #3.3-5 in the textbook, so you can check your answers.*

If $X$ is normally distributed with a mean of 6 and a variance of 25, find:

a.  $P(6 \leq X \leq 14)$
b.  $P(4 \leq X \leq 14)$
c.  $P(-4 \leq X \leq 0)$
d.  $P(X > 15)$
e.  $P(|X - 6| < 5)$
f.  $P(|X - 6| < 10)$
g.  $P(|X - 6| < 15)$
h.  $P(|X - 6| < 12.4)$
