---
title: "Lab 09"
subtitle: "MATH 361"
author: "Insert your name here"
date: "Insert due date here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's begin by loading the necessary packages. 

```{r, message = FALSE}
library(tidyverse)
```

# Tutorial

In this lab, you will practice simulating random variables from various distributions and estimating their means and variances. 

Recall an example from lecture. Let $X$ have the pdf $f(x) = 4x^3 , \ \ \ 0 < x < 1$.  If we want to simulate values from this distribution, we can find the cdf $Y = F(x)$ and its inverse $X = F^{-1}(y)$. We then utilize the fact that a cdf always follows a uniform (0,1) distribution. 

Note $f(x) = 4x^3 \implies Y =F(x) = \int_0^x4t^3dt=x^4$. Therefore, $F^{-1}(y) = X = Y^{1/4}$. So we can simulate values of $X$ by first simulating $Y$ values from the uniform distribution, and then using $F^{-1}(Y)$ to obtain values of $X$.

```{r}
#| message: false
#simulate 100,000 y values from the uniform distribution
y <- runif(100000)

#use the inverse cdf to compute x
x <- y^{1/4}

#plot a histogram of the random variable x
ggplot() +
   geom_histogram(aes(x = x), 
                  color = "white", boundary = 1) +
   theme_minimal()
```

Mathematically, we can find the mean and the variance of $X$ as follows:

$$\begin{aligned}
E(X) &= \int_0^1xf(x)dx \\
&= \int_0^1x4x^3dx \\
&= \int_0^14x^4dx \\
&= \frac{4}{5}x^5 \rvert_0^1 \\
&= \frac{4}{5}
\end{aligned}$$

$$\begin{aligned}
V(X) &= E(X^2) - [E(X)]^2 \\
&= \int_0^1x^2f(x)dx - \left[\frac{4}{5}\right]^2\\
&= \int_0^1x^24x^3dx - \left[\frac{4}{5}\right]^2\\
&= \int_0^14x^5dx - \left[\frac{4}{5}\right]^2\\
&= \frac{4}{6}x^6 \rvert_0^1 - \left[\frac{4}{5}\right]^2\\
&= \frac{4}{6} - \left[\frac{4}{5}\right]^2 \\
&= \frac{2}{125} = 0.02\overline{666}
\end{aligned}$$

We can check our work by computing the mean and variance of our simulate values in R. The built-in R functions are `mean()` and `var()`. Because these are simulated values, the mean and variance will not match exactly, but should be very close. 

```{r}
mean(x)
var(x)
```


# Tasks 

# Task 1

Simulate 100,000 values from the Cauchy distribution and plot a histogram of the distribution. Refer to #3 in your 5.1 Group Work. Recall that the cdf of a Cauchy distribution is $$F = \frac{1}{\pi}\left(arctan(x) + \frac{\pi}{2}\right) \ \ \ -\infty < x < \infty$$ *Hint: in R, the function `tan()` computes the tangent, and R recognizes the object `pi`*

```{r}
#simulate y from the uniform distribution

#compute x using the inverse

#plot histogram of x

```


# Task 2

Perform a simulation to estimate $Cov(V, W)$, where $V = X + Y$, $W = max(X, Y)$, and $X$, $Y$ are both distributed as $N(100, 15)$. To do this, 

1. Simulate 100,000 x-values using the `rnorm()` function. Recall that the `rnorm()` function requires that you specify `sd`, the standard deviation, not the variance. Run `?rnorm` in your console to pull up the help page.
2. Simulate 100,000 y-values using the `rnorm()` function
3. Create new objects `v` and `w` from the simulated `x` and `y` values. Note that the function `pmax()` will compute the pair-wise maximums for two vectors. 
4. Use the function `cov()` to compute the covariance between `v` and `w`

```{r}
#simulate 100,000 x-values


#simulate 100,000 y-values


#create v


#create w




#compute covariance 

```


# Task 3

Use R to compute the mean and variance of your simulated `x` and `y` values from Task 2. What do you expect these values to equal based on how $X$ and $Y$ are distributed? Are they good estimates of $E(X)$, $V(X)$, $E(Y)$, and $V(Y)$?

```{r}
#mean of x

#variance of x

#mean of y

#variance of y

```

> YOUR COMMENTS HERE

# Task 5

Exercise 5.1-6 on Homework 09 asks you to find the cdf of $X$ where $f(x) = 2x, 0 < x < 1$ and describe how you would simulate an observation of $X$. Using your work from parts a and b of that exercise, use R to:

1. simulate 100,000 values of $X$
2. plot a histogram
3. estimate the mean 
4. estimate the variance 

```{r}
#simulate y


#simulate x


#plot histogram


#compute mean of x


#compute variance of x



```

