---
title: "Lab 07"
subtitle: "MATH 361"
author: "Insert your name here"
date: "Insert due date here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

# Tutorial

This lab has you work with probabilities and percentiles from the normal distribution. Recall the following functions:

-   `pnorm(x, mean, sd)` - returns $F(x) = P(X \leq x)$ values for a normal random variable $X \sim N(\mu, \sigma^2)$ with parameter arguments `mean` = $\mu$, and `sd` = $\sigma$.

-   `qnorm(p, mean, sd)` - finds $x$ such that $P(X < x) = p$ for a normal random variable $X \sim N(\mu, \sigma^2)$ with parameter arguments `mean` = $\mu$, and `sd` = $\sigma$.

REMEMBER: the `_norm()` functions in R are defined in terms of the standard deviation $\sigma$, not the variance $\sigma^2$!

The following code finds $P(X < 1800)$ when $X$ is normally distributed with mean 1500 and variance 90000.

```{r}
pnorm(1800, 1500, sqrt(90000))
```

Note that `pnorm` and `qnorm` are inverse functions. The following code finds the percentile cutoff, $x$, such that $P(X \leq x)$ is 0.8413447. Based on our answer above, we should expect the following output to give us the score 1800.

```{r}
qnorm(0.8413447, 1500, sqrt(90000))
```

# Tasks

Note that both tasks are odd exercises from the HTZ textbook, so you can check your answers in the back of the book.

## Task 1

*\[Problem 3.3-1\]* If $Z$ is $N(0,1)$, find:

a.  $P(0.47 < Z < 2.13)$

```{r}

```

b.  $P(-0.97 \leq Z < 1.27)$

```{r}

```

c.  $P(Z > -1.56)$

```{r}

```

d.  $P(Z > 2.78)$

```{r}

```

e.  $P(|Z| < 1.96)$

```{r}

```

f.  $P(|Z| < 1)$

```{r}

```

g.  $P(|Z| < 2)$

```{r}

```

h.  $P(|Z| < 3)$

```{r}

```

i.  What rule do parts f - h demonstrate?

> YOUR ANSWER HERE

## Task 2

\[*Problem 3.3-3*\] If $Z$ is $N(0,1)$, find values of $c$ such that:

a.  $P(Z \geq c) = 0.025$

```{r}

```

b.  $P(|Z| \leq c) = 0.95$

```{r}

```

c.  $P(Z > c) = 0.05$

```{r}

```

d.  $P(|Z| \leq c) = 0.90$

```{r}

```

e.  Comment on how your answers in parts a - d relate to your answers in Task 1.

> YOUR ANSWER HERE
