---
title: "Lab 03 / 2.1 Group Work"
subtitle: "Intro to simulations"
author: "MATH 361"
date: "Spring 2024"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| label: load-packages
#| echo: false
#| message: false
library(tidyverse)
```

In this lab, you will actually conduct the simulations that you wrote psuedocode for in 2.1 Group Work.

Recall our random experiment of flipping a weighted coin 3 times, where each toss has a 70% probability of landing on heads. This could be used to model the \# of shots a 70% free throw shooter will make if they shoot 3 free throws (we are considering the shots to be independent).

We are interested in the following three random variables:

-   $X$ = the total number of heads (made shots)
-   $Y$ = whether or not there is a heads (made shot) on the 2nd flip
-   $Z$ = the number of heads (makes) minus the number of tails (misses)

## Task 1

Simulate 1 run of this experiment by creating three objects `toss1`, `toss2`, and `toss3`. You can make use of the function `rbernoulli()`, which takes two arguments:

-   n = number of runs of the experiment
-   p = probability of success

```{r}

```

## Task 2

Use the `toss1`, `toss2`, `toss3` objects to create your random variables `X`, `Y`, and `Z`.

```{r}

```

## Task 3

Adapt your code from Tasks 1 and 2 to simulate 100,000 values of `X`, `Y`, and `Z`

```{r}

```

## Task 4

The following code combines the vectors into a data frame, and creates summary tables and a visualization for the distribution of `X`. Remove `eval: false` so the output will be in your html.

```{r}
#| eval: false

# create data frame
sims <- data.frame(toss1, toss2, toss3, X, Y, Z)

# check dimensions - should be 100000 x 6
dim(sims)

# create frequency table
table(sims$X)

# create relative frequency table
proportions(table(sims$X))

# create visualization
ggplot(sims, aes(x = X)) +
  geom_bar()
```

Interpret these results in the context of the free throw shooting example.

> YOUR ANSWER HERE

## Task 5

Create the frequency table, relative frequency table, and bar plot for `Y`. Do the same for `Z`.

```{r}

```

Comprehension check: for each random variable, what should the counts in the frequency table add to? What should the relative frequencies add to?

> YOUR ANSWER HERE

## Task 6

Comment on the distributions of `Y` and `Z`. Do they fit your intuition? What do you notice about distributions of `X` and `Z`?

> YOUR ANSWER HERE

## Task 7

A great benefit of simulations is we can approximate the expected value of a random variable by calculating the mean of many simulated observations of that random variable. For example, we can get an approximate value of $E(X)$ by calculating `mean(X)`.

Compute the approximate expected value for `X`, `Y`, and `Z`.

```{r}

```

How do these compare to the true expected values you computed in Group Work 2.2?

> YOUR ANSWER HERE

## BONUS

Let's assume that these 3 free throws take place with \<1 second left on the clock, when the shooting team is down by 2. Use your simulations to find the approximate probability that:

-   The shooting team wins
-   The shooting team loses
-   The game goes to overtime
