---
title: "Lab 10 / Ch 6 R Activity"
subtitle: "MATH 361"
author: "Insert your name here"
date: "Insert due date here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this activity, we will start with a demo of why the formula for $s^2$ divides by $n-1$ instead of $n$. Then, you will conduct some exploratory data analysis on the the `gapminder` dataset as Lab 09.

## Load packages

We first need to load the `tidyverse`, `dslabs`, and `infer` packages using the `library()` command.

```{r}
#| message: false
library(tidyverse)
library(dslabs)
library(infer)
```

## Prepare the data

We can load a dataset called `gapminder` that contains data on 185 countries from the years 1960 to 2016. We are going to consider the distribution of the variable `life_expectancy`.

```{r}
data("gapminder", package = "dslabs") 
```

Let's first `filter` the data to keep only the year 2010.

```{r}
gapminder_2010 <- gapminder %>% 
  filter(year == 2010)
```

## Exploratory Data Analysis (Lab 12)

Let's create a histogram of `life_expectancy` across the 185 countries.

```{r}
ggplot(gapminder_2010) +
  geom_histogram(aes(x = life_expectancy), color = "white")  +
  xlab("Life expectancy") +
  ylab("Number of countries")
```

### Task 1

Describe the shape of the distribution. Comment on modality and skewness.

> YOUR ANSWER HERE

### Task 2

The following code will return a summary of a variable.

```{r}
summary(gapminder_2010$infant_mortality)
```

Use this `summary()` function to compute a summary of life expectancy in the data.

```{r}
#insert code here
```

Comment on what you notice.

> YOUR ANSWER HERE

Is the mean or the median life expectancy larger? Why is that?

> YOUR ANSWER HERE

### Task 3

Let's explore life expectancy by continent. Copy and paste the code above for creating a histogram, and then add a new layer (using the `+` sign) with the code `facet_wrap(~ continent, nrow = 2)`

```{r}
#insert code here
```

Comment on what you notice.

> YOUR ANSWER HERE

### Task 4

The following code is another way of visualizing life expectancy by continent.

```{r}
ggplot(gapminder_2010, aes(x = continent, y = life_expectancy)) +
  geom_boxplot() +
  labs(x = "Continent", y = "Life expectancy (years)",
       title = "Life expectancy by continent")
```

Comment on what you notice. Are any features of the data apparent here that weren't apparent in the faceted histograms?

> YOUR ANSWER HERE

### Task 5

Let's investigate the relationship between life expectancy and GDP per capita. Let's first create a new variable `gdp_per_capita` that scales a country's GDP by the number of people in the population.

```{r}
gapminder_2010 <- gapminder_2010 %>% 
  mutate(gdp_per_capita = gdp/population)
```

Create a scatterplot to investigate the relationship between life expectancy and GDP per capita. Hint: use `geom_point()` to create a scatterplot.

```{r}
#insert code here
```

Comment on what you notice.

> YOUR ANSWER HERE

### Task 6

Copy and paste the code for the scatterplot above, but add the arguments `color = continent` and `size = population` inside the `aes()` function.

```{r}
#insert code here
```

What did this do to the graph?

> YOUR ANSWER HERE

Comment on what you notice.

> YOUR ANSWER HERE

### BONUS

Utilize the `group_by` and `summarize()` functions to compute the mean and median life expectancy by continent. Hint: refer to how `group_by()` and `summarize()` were used in Lab 11 to compute the sample mean estimates for each replicate.

```{r}

```
