---
title: "Lab 08"
subtitle: "MATH 361"
author: "Insert your name here"
date: "Insert due date here"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tutorial

In this lab, you will use R to compute probabilities from the normal and t distributions and use what you've learned about linear combinations of normal random variables.

You will use the functions `pt()` and `qt()` to find probabilities and cutoff values from the t-distribution. These functions are analogous to the `p___()` and `q___()` functions you have used from other distributions. The t-distribution requires one parameter - the degrees of freedom; in R this argument is called `df`. You can pull up the help page for the t-distribution functions by typing `?pt()` into your console.

### Example 1

Find $P(44.4 \leq \overline{X} \leq 48.98)$ when $\overline{X}$ is the sample mean of $n = 16$ observations with $X \sim N(46.58,40.96)$

This implies $\overline{X} \sim N(46.58, 40.96/16)$

```{r}
pnorm(48.98, mean = 46.58, sd = sqrt(40.96/16)) - pnorm(44.42, mean = 46.58, sd = sqrt(40.96/16))
```

### Example 2

Find $t_{0.10}(34)$. This asks, what cutoff value gives 0.10 probability in the upper tail for a t-distribution with 34 degrees of freedom. Recall that functions starting with `q` allow us to find cutoff values for a given probability.

```{r}
qt(p = 0.1, df = 34, lower.tail = FALSE)
```

### Example 3

Find $P(−1.87 \leq T \leq 1.87)$ when $T \sim t(34)$

```{r}
pt(1.87, df = 34) - pt(-1.87, df = 34)
```

# Tasks

# Task 1

Let $X_1, X_2, ... X_{16}$ be a random sample from a normal distribution $N(77,25)$. Compute the following. *Hint: you should first determine how* $\overline{X}$ is distributed

a)  $P(77 \leq \overline{X} \leq 79.5)$
b)  $P(74.2 \leq \overline{X} \leq 78.4)$

```{r}

```

# Task 2

Find the 95th percentile for the standard normal distribution. Compare it to the 95th percentile for the t-distribution with df = 10, df = 1000, and df = 100,000.

```{r}
#95th percentile of standard normal

#95th percentile of t(10)

#95th percentile of t(1000)

#95th percentile of t(100000)

```

Which of these four distribution has the "fattest tails"?

> YOUR ANSWER HERE

Which two distributions are the most similar to one another?

> YOUR ANSWER HERE

# Task 3

Find $P(-2.06 < X < 2.06)$ when $X \sim t(25)$

```{r}

```

# Task 4

*\[HTZ 5.5-15, on group work\]*

Let the distribution of $T$ be $t(17)$. Find:

a.  $t_{0.01}(17)$
b.  $t_{0.95}(17)$
c.  $P(-1.74 < T < 1.74)$

```{r}

```
